---
title: "Homework 5"
output: html_document
---

## Question 8.1
Describe a situation or problem from your job, everyday life, current events, etc., for which a linear regression model would be appropriate. List some (up to 5) predictors that you might use.

**A:** In my job in the food delivery industry, a linear regression model can be used to predict the number of orders that may be made in a particular city, date, or time.

Some predictors that can be used:

- Square footage of the city
- Population of the city
- Day of the week
- Time of the day
- Weather (can be a boolean variable of whether it is raining)


## Question 8.2
Using crime data from http://www.statsci.org/data/general/uscrime.txt (file uscrime.txt, description at http://www.statsci.org/data/general/uscrime.html ), use regression (a useful R function is lm or glm) to predict the observed crime rate in a city with the following data:

M = 14.0,
So = 0,
Ed = 10.0,
Po1 = 12.0,
Po2 = 15.5,
LF = 0.640,
M.F = 94.0,
Pop = 150,
NW = 1.1,
U1 = 0.120,
U2 = 3.6,
Wealth = 3200,
Ineq = 20.1,
Prob = 0.04,
Time = 39.0

Show your model (factors used and their coefficients), the software output, and the quality of fit.
Note that because there are only 47 data points and 15 predictors, you’ll probably notice some overfitting. We’ll see ways of dealing with this sort of problem later in the course.


**A:** 

**General approach:**

- Scale the independent variables to ensure that all variables contribute equally to the linear regression model
- Since the dataset is rather small with only 47 observations, we will use a leave-one-out-cross-validation (LOOCV) method for the cross validation step
- With the trained model, we use the data points given in the question to predict the amount of crime in the city
- However, the model can be improved further as we see that some variables are not statistically significant at a 95% level in contributing to the model. Therefore, we can train the model again with only the statistically significant variables
- With this new model, we use the data points given in the question again to predict the amount of crime in the city, which gives a more appropriate estimate

Perform initialisation:
``` {r message = FALSE}
# Load required libraries
library(caret)
library(boot)
library(Hmisc)

# Set seed for reproducibility
set.seed(31)
```

First, we read the data into a dataframe and observe the general statistics of the dataset given.
``` {r}
# Read data into a dataframe
df <- read.table("C:/Users/Clayton/Documents/Gatech/ISYE 6501 Intro to Analytics Modelling/Module 6 Change Detection/Homework 3/uscrime.txt", header = TRUE)

describe(df)
```

Since we are using a linear regression model which can be sensitive to differing magnitudes of independent variables, we will need to scale the independent variables to ensure that each variable contributes to the model equally.
``` {r}
# Scale only the independent variables (columns 1-14)
df_scaled_features <- scale(df[1:(length(df)-1)])

# Combine the scaled independent variables and the unscaled dependent variable 
df_scaled <- merge(df_scaled_features, df[length(df)], by='row.names', all.x=TRUE)
df_scaled <- df_scaled[2:length(df_scaled)]
```

After scaling the features, we then perform an LOOCV method for cross validation since we have a small dataset of only 47 observations. This method ensures that we still have a large enough number of observations for model training before performing the validation. 

``` {r}
# Specify the cross-validation method
ctrl <- trainControl(method="LOOCV")

# Fit the linear regression model and use LOOCV to evaluate performance
model <- train(Crime ~ ., data=df_scaled, method="lm", trControl=ctrl)
```

This model has an RMSE of about `274` and a Multiple R-squared score of about `80%`. The factors and coefficients of the model are as shown below.

Here we observe that the Adjusted R-squared score (`~71%`) is lower than the Multiple R-squared score, which indicates that we may have included more predictors than necessary in this model. Therefore, we can retrain the model using only the statistically significant variables to ensure we do not include unnecessary variables.

``` {r}
summary(model)
model$results$RMSE
```

Using this current model, we can predict the amount of crime using the datapoints given in the question. Since we have scaled the independent variables for training, we will then need to scale this test datapoint using the scaling attributes of the training data before running the predictions.

From the prediction, we see that the predicted number of crime is about `155`, which is far lower than the smallest number of crime `342` in our training set. This could mean that the current model is not appropriate enough to predict this test datapoint and the prediction is based on extrapolating results from the linear regression model.

``` {r}
# Test data from the question
test_data = data.frame(
  M = 14.0,
  So = 0,
  Ed = 10.0,
  Po1 = 12.0,
  Po2 = 15.5,
  LF = 0.640,
  M.F = 94.0,
  Pop = 150,
  NW = 1.1,
  U1 = 0.120,
  U2 = 3.6,
  Wealth = 3200,
  Ineq = 20.1,
  Prob = 0.04,
  Time = 39.0
)

# Scale the test data using scaling attributes from training
test_scaled <- scale(test_data, center=attr(df_scaled_features, "scaled:center"), scale=attr(df_scaled_features, "scaled:scale"))

# Run the prediction
predict(model, newdata=test_scaled)
```

With this, we can build a better model using only the statistically significant independent variables for training. From our previous results, we use only variables that are statistically significant at a 95% level (M, Ed, Po1, U2, Ineq, Prob).

```{r}
# Train the model using only statistically significant variables
new_model <- train(Crime ~ M + Ed + Po1 + U2 + Ineq + Prob, data=df_scaled, method="lm", trControl=ctrl)
```

With this new model, we see that both RMSE and Adjusted R-squared score have improved from `274 -> 221` and `71% -> 73%` respectively. We can also observe that the difference between the Multiple R-squared score and the Adjusted R-squared score has reduced from `80% vs 71%` in the previous model to `77% vs 73%` in this new model, which can indicate that the number of predictors used in this model is more appropriate.

```{r}
summary(new_model)
new_model$results$RMSE
```

With the new model, the predicted number of crime is about `1,304`, which is within the range of the number of crime in the training set (`342 - 1,993`). Therefore, the prediction seems to be more appropriate with the new model.
```{r}
# Run the prediction using the new model, the model automatically picks out only the relevant variables for prediction
predict(new_model, newdata=test_scaled)
```