---
title: "Homework 5"
output: html_document
---

## Question 8.1
Describe a situation or problem from your job, everyday life, current events, etc., for which a linear regression model would be appropriate. List some (up to 5) predictors that you might use.

**A:** In my job in the food delivery industry, a linear regression model can be used to predict the number of orders that may be made in a particular city, date, or time.

Some predictors that can be used:

- Square footage of the city
- Population of the city
- Day of the week
- Time of the day
- Weather (can be a boolean variable of whether it is raining)


## Question 8.2
Using crime data from http://www.statsci.org/data/general/uscrime.txt (file uscrime.txt, description at
http://www.statsci.org/data/general/uscrime.html ), use regression (a useful R function is lm or glm) to
predict the observed crime rate in a city with the following data:
M = 14.0
So = 0
Ed = 10.0
Po1 = 12.0
Po2 = 15.5
LF = 0.640
M.F = 94.0
Pop = 150
NW = 1.1
U1 = 0.120
U2 = 3.6
Wealth = 3200
Ineq = 20.1
Prob = 0.04
Time = 39.0
Show your model (factors used and their coefficients), the software output, and the quality of fit.
Note that because there are only 47 data points and 15 predictors, you’ll probably notice some
overfitting. We’ll see ways of dealing with this sort of problem later in the course.


**A:** 

**General approach:**

- Use triple exponential smoothing model to estimate for beta (trend) and gamma (seasonality), the beta estimate measures whether the annual temperatures have trended up or down over the years
- Use CUSUM method on the average annual seasonality factors to identify whether seasonality factors has increased in the later years, which would indicate that unofficial end of summer has gotten later
- From the findings, we see that beta estimate is 0, which indicates that the annual temperature trend has remained flat, and the CUSUM on average annual seasonality factors have largely remained the same, indicating that the seasonality factors have also remained the same over the years
- Therefore, we can conclude that unofficial end of summer has not gotten later over the past 20 years

Perform initialisation:
``` {r message = FALSE}
# Load required libraries
library(qcc)

# Set seed for reproducibility
set.seed(31)
```

First, we read the data into a dataframe, convert the dataframe into a vector, then finally to a time-series object for the HoltWinters model
``` {r}
# Read data into df
df <- read.table("C:/Users/Clayton/Documents/Gatech/ISYE 6501 Intro to Analytics Modelling/Module 7 Time Series Models/Homework 4/temps.txt", header = TRUE)

# Convert dataframe into a vector  
df_vector <- as.vector(unlist(df[, 2:21]))

# Convert vector to time-series object for the HoltWinters model
df_length <- length(df[, 1])
temp_ts <- ts(df_vector, start=1996, frequency=df_length)
```

Plot the time-series data for visualisation
``` {r}
plot(temp_ts)
```

From the chart, the trend of the temperature data seem to be rather flat visually. We can verify this using the exponential smoothing model.

To analyse whether the unofficial end of summer has gotten later over the years, we are generally interested in 2 factors: the beta (or trend) estimate and the gamma (or seasonality) estimate.
Therefore, we will be using the Triple Exponential Smoothing approach to determine the trend and seasonality estimates.
For the Seasonality estimate, we will use the 'Additive' approach as we will generally expect temperature differences between summer and fall for each year to be roughly consistent.

``` {r}
es_triple_additive_model <- HoltWinters(temp_ts, seasonal='additive')
```

``` {r}
es_triple_additive_model$beta
es_triple_additive_model$coefficients[2]
```

Here we see that beta (or trend estimate) and beta coefficient is 0 or near 0, which means that across the past 20 years, the trend of temperatures from July to October has remained flat.
However, this is not sufficient to conclude that the unofficial end of summer over the past 20 years has remained the same.
For this, we will need to observe the seasonality factors to determine whether unofficial end of summer has gotten later.

The intuition here is that if unofficial end of summer has gotten later, we will expect the average annual seasonality factor to increase over the years, since we expect more days of summer between July and October in the later years.
We will use the CUSUM approach on the average annual seasonality factor to determine whether the annual seasonality factor has consistently increased above 5 standard errors in the later years.

``` {r}
# Extract the seasonality factors from the Triple Exponential Smoothing model to a matrix
season_matrix <- matrix(es_triple_additive_model$fitted[, 4], ncol=df_length)

# Get the average annual seasonality factors for all years
avg_seasonality <- data.frame(rowMeans(season_matrix, n=19))
```

For the CUSUM method, we will use the mean annual seasonality factor as the center (C) and the threshold (T) to be a standard 5x of the standard error.
```{r}
# Get the mean annual seasonality factor for all years for CUSUM center (C)
mean_seasonality <- mean(avg_seasonality[, 1])

# Use the mean annual seasonality factor as the center (C) and the threshold (T) to be a standard 5x of the standard error
seasonality_cusum <- cusum(avg_seasonality, center=mean_seasonality, decision.interval=5)
```

From the CUSUM chart, we see that the average annual seasonality factor for most of the remaining 19 years has remained within 5 standard errors.
The only exception is in the year 2010, when the seasonality factor increased above 5 standard errors; however, from the chart, this is a one-off observation and the seasonality factor subsequently reverted to the mean afterwards.
Therefore, we can conclude that the **seasonality factors across the years have largely remained the same** and **unofficial end of summer has not gotten later in the past 20 years**.
