---
title: "Homework 9"
output: html_document
---

# Question 12.1
Describe a situation or problem from your job, everyday life, current events, etc., for which a design of experiments approach would be appropriate.

**A:** Taking the example of a food delivery app, a design of experiments approach can be used to identify which version of food item recommendations is most useful in driving more orders. 

For example, we can test 2 different variations of recommendations: 1 using the most popular food items on the app and another 1 using similar food items that a customer has purchased in the past. 

Based on the 2 versions, we can then use a statistical approach to determine which version is better at driving more clicks (and therefore orders) and to what degree, as well as its significance level.

# Question 12.2
To determine the value of 10 different yes/no features to the market value of a house (large yard, solar roof, etc.), a real estate agent plans to survey 50 potential buyers, showing a fictitious house with different combinations of features. To reduce the survey size, the agent wants to show just 16 fictitious houses. Use R’s FrF2 function (in the FrF2 package) to find a fractional factorial design for this experiment: what set of features should each of the 16 fictitious houses have? Note: the output of FrF2 is “1” (include) or “-1” (don’t include) for each feature.

**A:**
Initialisation 
``` {r, message=FALSE, warning=FALSE}
library(FrF2)
```

Using a fictitious set of 10 features: `Large Yard`, `Solar Roof`, `Large Garage`, `>2 bathrooms`, `>3 bedrooms`, `Multiple floors`, `Near public transport`, `Quiet neighbourhood`, `Marble flooring`, `<10 years old`; we use FrF2 package to create the set of 16 fictitious houses with the following set of features:
``` {r}
FrF2(
  16, 
  factor.names = c("large_yard", "solar_roof", "large_garage", "more_than_2_bathrooms", "more_than_3_bedrooms", "multiple_floors", "near_public_transport", "quiet_neighbourhood", "marble_flooring", "less_than_10_years_old"), 
  default.levels = c("yes", "no")
)
```

# Question 13.1
For each of the following distributions, give an example of data that you would expect to follow this distribution (besides the examples already discussed in class).

#### a. Binomial 

**A:** Number of customers making a complaint when serving N customers

#### b. Geometric 

**A:** Number of customers served before receiving a customer complaint 

#### c. Poisson 

**A:** Average number of customers arriving in a 1-hour period

#### d. Exponential 

**A:** Average time between each customer arrival

#### e. Weibull

**A:** Average time between each customer complaint

# Question 13.2
In this problem you can simulate a simplified airport security system at a busy airport. Passengers arrive according to a Poisson distribution with λ1 = 5 per minute (i.e., mean interarrival rate 1 = 0.2 minutes) to the ID/boarding-pass check queue, where there are several servers who each have exponential service time with mean rate 2 = 0.75 minutes. [Hint: model them as one block that has more than one resource.] After that, the passengers are assigned to the shortest of the several personal-check queues, where they go through the personal scanner (time is uniformly distributed between 0.5 minutes and 1 minute). Use the Arena software (PC users) or Python with SimPy (PC or Mac users) to build a simulation of the system, and then vary the number of ID/boarding-pass checkers and personal-check queues to determine how many are needed to keep average wait times below 15 minutes. [If you’re using SimPy, or if you have access to a non-student version of Arena, you can use λ1 = 50 to simulate a busier airport.]

**A:**
Installing python packages to enable running the simulation on R.
``` {r, message=FALSE, warning=FALSE}
library(reticulate)
py_install(packages = c("simpy", "numpy"))
```

To create the simulation, we first need to set up the constants that are given in the question to define the premise of this simulation. All the constants are given in the question above
``` {python, message=FALSE, warning=FALSE} 
import random
from numpy import random
import simpy

# Set seed
seed = 31

# Constants from the question
passenger_lam = 5 #passenger arrival lambda = 5
queue_scale = 1 / 0.75 #scale is the inverse of rate (which is 0.75 from the question)
scanner_min = 0.5 #scanner minimum process time
scanner_max = 1.0 #scanner maximum process time
```

Here, our aim is to find the number of ID/boarding pass check queues and the number of personal scanners in the airport needed to ensure that average waiting times are kept within 15 minutes. From my simulations, I found that minimally the airport will need 3 ID/boarding pass check queues and 2 personal scanners to keep waiting times within 15 minutes. We can set the simulation time here as well, which I have used 2 hours as the simulation time.
``` {python, message=FALSE, warning=FALSE} 
# Parameters to tweak
sim_time = 120 #Number of minutes to run the simulation, let's use 2 hours for this
num_queues = 3
num_scanners = 2
```

Here, we define the airport class with 2 resources: ID/boarding pass checkers & personal scanners. The number of checkers and scanners are from the above parameters.

In the airport class, it also has 2 functions: the check function for ID/boarding pass & the scan function for personal scans. As mentioned in the question, the ID/boarding pass checking follows an exponential distribution while the personal scanner follows a uniform distribution; both functions are defined with their respective distributions.

In the passenger function, we run every passenger through the checking and scanning process. Here, we start measuring the passenger wait times from when they arrive at the airport till they complete the entire checking & scanning process. We then get the wait times of all passengers and find the average waiting time for each configuration of checkers & scanners.

In the setup function, we create the airport class and define the passenger arrivals. As mentioned in the question, passenger arrivals follow a poisson distribution so we will model it as such. With every passenger arrival, we will subject the passenger to the checking and scanning process.

In the simulation, the logs are commented out for brevity.
``` {python, message=FALSE, warning=FALSE} 
class Airport(object):
    '''
    The airport has 2 resources: the ID/boarding pass checking queue and the personal scanners
    These 2 resources are defined in the parameters
    '''
    def __init__(self, env, queue_scale, scanner_min, scanner_max, num_queues, num_scanners):
        self.env = env
        self.checkers = simpy.Resource(env, num_queues)
        self.scanners = simpy.Resource(env, num_scanners)
        self.queue_scale = queue_scale
        self.scanner_min = scanner_min
        self.scanner_max = scanner_max 
        
    def check(self, passenger):
        '''
        The ID/Boarding pass checking process based on exponential distribution of mean rate 0.75 minutes
        '''
        queue_time = random.exponential(scale=1/self.queue_scale, size=1)[0]
        yield self.env.timeout(queue_time)
        
    def scan(self, passenger):        
        '''
        The personal scanner process based on uniform distribution between 0.5 to 1 minute
        '''
        scan_time = random.uniform(self.scanner_min, self.scanner_max)
        yield self.env.timeout(scan_time)
        
def passenger(env, name, airport, wait_time_list):
    '''
    Passenger that arrives at the airport. 
    The waiting time starts when they arrive and ends when they complete the final scan.
    Here they have to go through the ID/boarding pass check and the personal scan
    '''
    # print(f'{name} arrives at the airport at {env.now}.')
    start_time = env.now

    with airport.checkers.request() as request:
        yield request 
        
        # print(f'{name} enters the ID/boarding pass checking queue at {env.now}.')
        yield env.process(airport.check(name))
        
    with airport.scanners.request() as request:
        yield request 
        
        # print(f'{name} enters the personal scanner at {env.now}.')
        yield env.process(airport.scan(name))
        
        # print(f'{name} completes the process at {env.now}.')
        end_time = env.now

    wait_time_list.append(end_time - start_time) #Get the total wait time of each passenger at the airport

def setup(env, passenger_lam, queue_scale, scanner_min, scanner_max, num_queues, num_scanners, wait_time_list):
    '''
    Create the airport object, with passengers arriving at the airport every 0.2 minutes
    Passenger arrival rate is based on inverse of poisson distribution with lambda of 5
    '''
    airport = Airport(env, queue_scale, scanner_min, scanner_max, num_queues, num_scanners)

    passenger_count = 0

    while True:
        passenger_arrival = random.poisson(lam=passenger_lam, size=1)[0]
        if passenger_arrival == 0:
            passenger_arrival = 0.0001
        yield env.timeout(1/passenger_arrival)
        passenger_count += 1
        env.process(passenger(env, f'Passenger {passenger_count}', airport, wait_time_list))
```

With the setup completed, we can finally run the simulation and identify the average waiting time of passengers when we have 3 ID/boarding pass checking queues & 2 personal scanners at the airport. The average waiting time is about `4.23` minutes.
``` {python, message=FALSE, warning=FALSE} 
# Simulation
random.seed(seed)         
env = simpy.Environment()
wait_time_list = []

env.process(setup(env, passenger_lam, queue_scale, scanner_min, scanner_max, num_queues, num_scanners, wait_time_list))

env.run(until=sim_time)
print(f'The average waiting time for each passenger is {round(sum(wait_time_list) / len(wait_time_list), 2)} minutes')
```

Extending the simulation further with more passenger arrivals as mentioned in the question, we now define the passenger arrival lambda to be 50 instead of 5.

With more passenger arrivals, we will need 29 ID/boarding pass checking queues & 29 personal scanners at the airport to ensure average waiting times are kept under 15 minutes. With this setup, the average waiting time is about `14.93` minutes.
``` {python, message=FALSE, warning=FALSE} 
# Set seed
seed = 31

# Constants from the question
passenger_lam = 50 #increase lambda to 50
queue_scale = 1 / 0.75 #scale is the inverse of rate (which is 0.75 from the question)
scanner_min = 0.5 #scanner minimum process time
scanner_max = 1.0 #scanner maximum process time

# Parameters to tweak
sim_time = 120 #Number of minutes to run the simulation, let's use 2 hours for this
num_queues = 29
num_scanners = 29

# Simulation
random.seed(seed)         
env = simpy.Environment()
wait_time_list = []

env.process(setup(env, passenger_lam, queue_scale, scanner_min, scanner_max, num_queues, num_scanners, wait_time_list))

env.run(until=sim_time)
print(f'The average waiting time for each passenger is {round(sum(wait_time_list) / len(wait_time_list), 2)} minutes')
```