---
title: "Homework 9"
output: html_document
---

# Question 12.1
Describe a situation or problem from your job, everyday life, current events, etc., for which a design of experiments approach would be appropriate.

**A:** Taking the example of a food delivery app, a design of experiments approach can be used to identify which version of food item recommendations is most useful in driving more orders. 

For example, we can test 2 different variations of recommendations: 1 using the most popular food items on the app and another 1 using similar food items that a customer has purchased in the past. 

Based on the 2 versions, we can then use a statistical approach to determine which version is better at driving more clicks (and therefore orders) and to what degree, as well as its significance level.

# Question 12.2
To determine the value of 10 different yes/no features to the market value of a house (large yard, solar roof, etc.), a real estate agent plans to survey 50 potential buyers, showing a fictitious house with different combinations of features. To reduce the survey size, the agent wants to show just 16 fictitious houses. Use R’s FrF2 function (in the FrF2 package) to find a fractional factorial design for this experiment: what set of features should each of the 16 fictitious houses have? Note: the output of FrF2 is “1” (include) or “-1” (don’t include) for each feature.

**A:**
Initialisation 
``` {r, message=FALSE, warning=FALSE}
library(FrF2)
```

Using a fictitious set of 10 features: `Large Yard`, `Solar Roof`, `Large Garage`, `>2 bathrooms`, `>3 bedrooms`, `Multiple floors`, `Near public transport`, `Quiet neighbourhood`, `Marble flooring`, `<10 years old`; we use FrF2 package to create the set of 16 fictitious houses with the following set of features:
``` {r}
FrF2(
  16, 
  factor.names = c("large_yard", "solar_roof", "large_garage", "more_than_2_bathrooms", "more_than_3_bedrooms", "multiple_floors", "near_public_transport", "quiet_neighbourhood", "marble_flooring", "less_than_10_years_old"), 
  default.levels = c("yes", "no")
)
```

# Question 13.1
For each of the following distributions, give an example of data that you would expect to follow this distribution (besides the examples already discussed in class).

#### a. Binomial 

**A:** Number of customers making a complaint when serving N customers

#### b. Geometric 

**A:** Number of customers served before receiving a customer complaint 

#### c. Poisson 

**A:** Average number of customers arriving in a 1-hour period

#### d. Exponential 

**A:** Average time between each customer arrival

#### e. Weibull

**A:** Average time between each customer complaint

# Question 13.2
In this problem you can simulate a simplified airport security system at a busy airport. Passengers arrive according to a Poisson distribution with λ1 = 5 per minute (i.e., mean interarrival rate 1 = 0.2 minutes) to the ID/boarding-pass check queue, where there are several servers who each have exponential service time with mean rate 2 = 0.75 minutes. [Hint: model them as one block that has more than one resource.] After that, the passengers are assigned to the shortest of the several personal-check queues, where they go through the personal scanner (time is uniformly distributed between 0.5 minutes and 1 minute). Use the Arena software (PC users) or Python with SimPy (PC or Mac users) to build a simulation of the system, and then vary the number of ID/boarding-pass checkers and personal-check queues to determine how many are needed to keep average wait times below 15 minutes. [If you’re using SimPy, or if you have access to a non-student version of Arena, you can use λ1 = 50 to simulate a busier airport.]
